---
title: "Report on analyzing the current state of the legal cartography of watercourses in France"
format: docx
editor: visual
execute-dir: project
execute:
  echo: false
  warning: false
---

```{r load-data}
library(data.table)
library(ggplot2)
library(stringr)
library(targets)

tar_load(metadata_nets_formatted)
tar_load(ddtnets_bvinters)
tar_load(ddt_nets_colnas)

knit_print.integer = function(x, ...) {
  prettyNum(x, big.mark=",")
}
```

# Metadata

```{r summarize metadata}
ddtnets_n_reported <- as.integer(sum(metadata_nets_formatted$nlines_reported))
ddtnets_n <- ddtnets_bvinters[!duplicated(UID_CE), .N]
ddtorig_n_online <- metadata_nets_formatted[
  !duplicated(local_url) &
    (data_orig %in% c('GEOIDE', 'Autre site')), 
  .N]
ddtorig_n_ddt <- metadata_nets_formatted[
  !duplicated(local_url) & 
    (data_orig %in% c('DDT', 'Données procurées à Hervé Pella par DDT')), 
  .N]


ddtnets_attris <- metadata_nets_formatted[!duplicated(local_url),
                                            str_split_1(attri_names, ';'),
                                            by='dep_code'] %>%
  .[V1 != '',]
ddtnets_attris_n <- ddtnets_attris[, .N, by='dep_code']

#Check matching in attribute names among departments (removing spaces and caps)
attrimatches_pairwise_avg  <- as.data.table(
  t(combn(ddtnets_attris[, paste(dep_code, 
                                 str_to_lower(
                                   gsub('(^\\s)|(\\s$)', '', V1, perl=TRUE)
                                 ), 
                                 sep=";")],
          2))) %>%
  .[, c('dep_code_1' , 'attri_name_1') := tstrsplit(V1, ';')] %>%
  .[, c('dep_code_2' , 'attri_name_2') := tstrsplit(V2, ';')] %>%
  .[dep_code_1 != dep_code_2] %>%
  .[, -c('V1', 'V2')] %>%
  .[, sum(attri_name_1==attri_name_2, na.rm=T), 
    by=c('dep_code_1', 'dep_code_2')] %>%
  .[, mean(V1)]
```

There are `r length(unique(metadata_nets_formatted$local_url))` unique networks for mainland France, for a total of `r ddtnets_n_reported` individual watercourse segments. After removing invalid and duplicate geometries, there remain `r ddtnets_n` individual segments in the compiled dataset.

We obtained `r ddtorig_n_online` datasets directly online and `r ddtorig_n_ddt` from requests to the DDTs.

Most layers that are available online are regularly updated following yearly, or semestral meetings to integrate re-classification requests, as shown by the distribution of "revision date" provided in the online metadata associated with the layers (n=`r metadata_nets_formatted[!is.na(date_rev), .N]`/`r metadata_nets_formatted[, .N]` departments):

```{r date_rev plot}
ggplot(metadata_nets_formatted, aes(x=as.Date(date_rev))) +
  geom_histogram() +
  scale_x_date(
    name='Reported revision date',
    date_labels=str_wrap("%m-%Y", 5), date_breaks  ="6 month",
    expand=c(0,0)) + 
  scale_y_continuous(
    name='Number of departments',
    expand=c(0,0)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

The attributes associated with the data were varied, with `r`length(unique(ddtnets_attris\$V1))`unique attribute names across all networks. On average, the networks included`r mean(ddtnets_attris_n)`attributes (min-max:`r min(ddtnets_attris_n)`-`rmax(ddtnets_attris_n`). These attributes were more or less common to the various networks. After removing formatting differences (different uses of caps and spaces), any two pair of network had`r `round(attrimatches_pairwise_avg, 2)` attributes in common, on average.

In its guidelines to DDT in 2015, the ONEMA proposed that a minimum set of standardized fields be used by DDTs:

-   TYPE_ECOUL - Type d'écoulement: Cours d'eau, Indéterminé, Non cours d'eau, Inexistant

-   NAT_IDENT - Méthode d'identification de l'écoulement: Analyse cartographique, Terrain, Décision de justice

-   DATE_IDENT - Date de l'identification du type d'écoulement par les experts concernés

-   ORIG_MODIF - Source de la modification, de la suppression du tronçon BD TOPO®, ou de l'ajout d'un nouveau tronçon: e.g., Levé GPS BDCarthage SCAN25 BDORTHO BDPARCELLAIRE CARTE d'étatmajor Cadastre napoléonien Atlas ancien Autre

-   CODE_HYDRO - Code générique du cours d'eau (identifiant pour la codification hydrographique) A reporter depuis la BD CARTHAGE® si le tronçon correspond à un cours d'eau codifié

-   AUTEUR - Service qui a effectué la modification (ex : SPE55)

The following attributes are the most common in the networks:

```{r attri_names plot}
#Check which departments have an attribute that fits with the ONEMA recommendations
#Counting both without and with manual name standardization
ddtnets_standardized_attris <- metadata_nets_formatted[
  , c("dep_code", "type_stand_name","regime_name",
      "nat_id_name", "orig_mo_name", "date_id_name"),
  with=F] %>%
  setnames(
    c("type_stand_name", "regime_name", "nat_id_name","orig_mo_name", "date_id_name"),
    c("type_ecoul", "regime", "nat_ident", "orig_modif", "date_ident")) %>%
  melt(id.vars='dep_code') %>%
  .[, list(
    attri_name= str_split_1(
      str_to_lower(gsub('(^\\s)|(\\s$)', '', value, perl=TRUE)),
      ';')
  ),
  by=c('dep_code','variable')]


onema_attris <- c('type_ecoul', 'nat_ident', 'date_ident', 'orig_modif',
                  'code_hydro', 'auteur', 'comm')
bdtopo_attris <- c('id_loc', 'id', 'prec_plani', 'prec_alti', 'artif', 'fictif', 
                   'franchisst', 'nom', 'pos_sol', 'regime', 'z_ini', 'z_fin', 
                   'prec_p_m', 'prec_a_m', 'artif_m', 'fictif_m', 'franchis_m', 
                   'nom_m', 'pos_sol_m', 'regime_m', 'z_ini_m', 'z_fin_m')

attris_all <- ddtnets_attris[, list(
  attri_name= str_to_lower(gsub('(^\\s)|(\\s$)', '', V1, perl=TRUE))), 
  by='dep_code'] %>%
  merge(ddtnets_standardized_attris, 
        by=c('attri_name', 'dep_code'), all.x=T) %>%
  .[, is_standardized := fifelse((variable == attri_name) | is.na(variable), 
                              'no', 'yes')] %>%
  .[, standardized_name := variable] %>%
  .[is.na(variable), standardized_name := attri_name] %>%
  .[standardized_name %in% onema_attris, attri_orig := 'ONEMA'] %>%
  .[standardized_name %in% bdtopo_attris, attri_orig := 'BDTOPO'] %>%
  .[is.na(attri_orig), attri_orig:='Other'] %>%
  .[, n := .N, by=c('standardized_name', 'is_standardized')] %>%
  .[, n_total := .N, by=standardized_name]


ggplot(attris_all[n_total>20,], aes(x=reorder(standardized_name,-n_total), 
                              fill=attri_orig, 
                              alpha=is_standardized)
       ) +
  geom_histogram(stat='count', 
                 position = position_stack(reverse = TRUE)) +
  scale_x_discrete(name='Attribute name') +
  scale_y_continuous(
    name='Number of departments',
    limits=c(0, length(unique(ddtnets_attris$dep_code))),
    expand=c(0,0)) +
  scale_fill_discrete(name='Origin of the attribute') +
  scale_alpha_discrete(name= str_wrap('After manual name standardization', 20),
                       range = c(1, 0.6)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

DDTs were supposed to use the BD TOPO (TRONCON_COURS_EAU), v 15.1 (2015) as the hydrographic foundation to build their cartography (to be completed -- look for individual departments for which this information is still not available). Double check website and only cartography + search the internet?

```{r attri_nas plot}
#Compute percentage of NAs in cols of all DDTs
ddt_nets_colnas[, perc_null := 100*(n_null/nrows)]
attris_all_colnas <- merge(ddt_nets_colnas, 
                           attris_all[n_total>20, -c('attri_name')], 
                           by.x=c('attri_name', 'dep_code'),
                           by.y=c('standardized_name', 'dep_code'))

ggplot(attris_all_colnas, 
       aes(x=reorder(attri_name, -n_total), y=perc_null, fill=attri_orig)) +
  geom_boxplot(alpha=0.75) +
  scale_x_discrete(name='Attribute name') +
  scale_y_continuous(
    name='Percentage of records without a value',
    #limits=c(0, length(unique(ddtnets_attris$dep_code))),
    expand=c(0,0)) +
  scale_fill_discrete(name='Origin of the attribute') +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=45, hjust=1))


```

Even for a given attribute, there was substantial variability. For the attribute describing the status of the hydrographic segments under water law, there are 5 possible categories: watercourse, not watercourse, undetermined, inexistant and outside of the department.

```{r type_stand_cats}

type_ecoul_cats <- metadata_nets_formatted[
  , str_to_lower(gsub('(^\\s)|(\\s$)', '', 
                      str_split_1(type_stand_cats, ';'), 
                      perl=TRUE)
  ),
  by='dep_code']
type_ecoul_cats_N <- type_ecoul_cats[, .N, by=dep_code]
```

Nonetheless, there were a total of `r length(unique(type_ecoul_cats$V1))` unique categories for this attribute across the networks (including typos), for an average of `r round(mean(type_ecoul_cats_N$N), 1)` per network. The hydrographic segments that were considered not to qualify as a watercourse were provided for `r round(metadata_nets_formatted[, nrow(.SD[nce_recat!="Pas de catégorie correspondante",])/.N])` of departments.

# Drainage density

